```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, magrittr, gt, gtExtras)
```

# Guatemala's trade flows with four of its main commercial partners

Almost all of this code is an adaptation of [\@leeolney3](https://twitter.com/leeolney3) [#TidyTuesday](https://twitter.com/hashtag/TidyTuesday?src=hashtag_click) submission, which is itself an adaptation of [\@BjnNowak](https://twitter.com/BjnNowak) tutorials. For me it was a great opportunity for getting to know more of the `gt` and `gtExtras` packages.

## (a) Import data

The data set used here was made with the API wrapper for the United Nations Comtrade Database, `comtradr`. See <https://git.io/JSDlL> for details on the data extraction process.

```{r}
data_raw <- 
  read_csv("https://raw.githubusercontent.com/gafnts/tidy-tuesday/main/2022/01%20%E2%80%94%20Bring%20your%20own%20data/Data/data.csv") %>% 
  select(partner, year, trade_flow, trade_value_usd) %>% 
  pivot_wider(names_from = trade_flow, values_from = trade_value_usd) %>% 
  janitor::clean_names() %>% 
  arrange(partner, year) %>% 
  mutate(imports = round(import/1e6, 2), 
         exports = round(export/1e6, 2)) %>% 
  select(partner, year, imports, exports)
```

## (b) Data wrangling

```{r}
abs <-
  data_raw %>%
  pivot_longer(3:4, values_to = "absolute")

rel <-
  data_raw %>%
  mutate(imports = (imports / lag(imports) - 1),
         exports = (exports / lag(exports) - 1)) %>%
  replace(is.na(.), 0) %>%
  pivot_longer(3:4, values_to = "relative")

data <-
  abs %>%
  left_join(rel) %>%
  mutate(
    rank = case_when(
      partner == "USA" ~ 1,
      partner == "China" ~ 2,
      partner == "Mexico" ~ 3,
      partner == "El Salvador" ~ 4
    )
  )
```

## (c) Plots

```{r}
abs_fun <- function(data) {
  trend <- ggplot(data, aes(x = year, y = absolute, color = fct_rev(name))) +
    geom_line(size = 3.5) +
    scale_color_manual(values = c("#323E40","#F2A922")) +
    coord_cartesian(clip = "off") +
    cowplot::theme_minimal_grid(35) +
    theme(legend.position = "none",
          axis.title = element_blank())
  return(trend)
}

abs_plots <-
  data %>%
  group_by(rank) %>%
  nest() %>%
  mutate(gg = purrr::map(data, abs_fun)) %>%
  select(rank = rank, gg) %>%
  arrange(rank)
```

```{r}
rel_fun <- function(data) {
  trend <- ggplot(data, aes(
    x = year,
    y = relative,
    color = fct_rev(name)
  )) +
    geom_line(size = 3.5) +
    scale_color_manual(values = c("#323E40","#F2A922")) +
    scale_y_continuous(
      label = scales::percent
    ) +
    coord_cartesian(clip = "off") +
    cowplot::theme_minimal_grid(35) +
    theme(legend.position = "none",
          axis.title = element_blank())
  return(trend)
}

rel_plots = data %>%
  group_by(rank) %>%
  nest() %>%
  mutate(gg=purrr::map(data, rel_fun)) %>%
  select(rank = rank, gg) %>%
  arrange(rank)
```

## (d) Table

```{r}
base <- 
  data_raw %>%
  filter(year == 2020) %>%
  mutate(
    fl = case_when(
      partner == "USA" ~ "https://hatscripts.github.io/circle-flags/flags/us.svg",
      partner == "China" ~ "https://hatscripts.github.io/circle-flags/flags/cn.svg",
      partner == "Mexico" ~ "https://hatscripts.github.io/circle-flags/flags/mx.svg",
      partner == "El Salvador" ~ "https://hatscripts.github.io/circle-flags/flags/sv.svg"
    )
  ) %>%
  select(fl, partner, imports, exports) 
```

```{r}
table <-
  base %>%
  arrange(desc(imports)) %>%
  mutate(absolute = NA,
         relative = NA) %>%
  gt() %>%
  gt_theme_nytimes() %>%
  # Add country flag
  gt_img_rows(fl) %>%
  # Add line plot
  gt::text_transform(
    locations = cells_body(columns = absolute),
    fn = function(x) {
      purrr::map(abs_plots$gg,
                 gt::ggplot_image,
                 height = px(100),
                 aspect_ratio = 1.7)
    }
  ) %>%
  gt::text_transform(
    locations = cells_body(columns = relative),
    fn = function(x) {
      purrr::map(rel_plots$gg,
                 gt::ggplot_image,
                 height = px(100),
                 aspect_ratio = 1.7)
    }
  ) %>%
  # Spanner
  tab_spanner(label = "2020", columns = imports:exports) %>%
  tab_spanner(label = "1995-2020", columns = absolute:relative) %>%
  # Column labels
  cols_label(
    imports = html("<span style = 'color:#323E40;'>imports</span>"),
    exports = html("<span style = 'color:#F2A922;'>exports</span>"),
    absolute = html("Trend"),
    relative = html("Relative change"),
    fl = html(""),
    partner = html("Country")
  ) %>%
  tab_header(
    title = "Guatemala's trade flows with four of its main commercial partners",
    subtitle = md(
      "In millions of current US dollars (calculated using an average annual exchange rate)"
    )
  ) %>% 
  fmt_number(column = c("imports", "exports"), use_seps = TRUE) %>%
  tab_options(heading.title.font.size = "30px",
             heading.padding = "10px",
             footnotes.padding = "10px",
             source_notes.font.size = "10px",
             column_labels.font.weight = "bold",
             source_notes.padding = "10px") %>%
  cols_align(columns = c(1, 2, 3, 4, 5, 6),
             align = "center")  %>%
  tab_source_note(source_note = gt::html("Source: United Nations Comtrade Database")) %>% 
  opt_table_font(font = c(google_font(name = "Noto Sans"),
                          default_fonts()))
```

```{r}
table %>%
  gtsave(
    "trade_flows.png",
    path = paste0(here::here(), "/2022/01 — Bring your own data")
  )
```
